### 2009-09-02 doc/tech_doc/janitor/tex_concept/concept.tex
Making Janitor documentation compile in usual (almost) way

 The main language for the implementation of the functionality of the
 dynamic runtime environment functionality is Perl. And it is solely required (exceptions are the
 integration with the Grid Manager and the Web server) for the Janitor. In
 the pre-web-service implementation the Catalog remains a static web page.
 The Perl code is split into multiple modules as depicted in Figure~\ref{fig:janitorDependencies}. 
 The modules can be separated into two
 functional groups. One addresses the retrieval of information from
 the Catalog's RDF file in the left major branch of the figure. The other
 addresses the process of fetching and installing the packages.

 \begin{landscape}
\begin{figure}[!h]
\vspace{4cm}
 \begin{center} 
%    \resizebox{24cm}{!}{\input{images/dependencies}}
    \includegraphics[width=24cm]{images/dependencies.png}
    \mycaption{Modules of the Janitor and their dependencies}{}
 \end{center} 
\vfill
 \label{fig:janitorDependencies} 
\end{figure}
 \end{landscape}

@@

\section{States of runtime environments}
A major motivation for the managed, manual initiation of dynamic RE
installation is the subsequent manual verification of the installed
packages -- prior to their use in production.
With an automation of the installation, the verification of
that process shall be performed externally to that process.  At this
time, only the automation of the installation has been implemented.
To reflect the progress the external verification has made, REs
are said to be in states. The current implementation lists
installable REs aside the installed REs in the grid information system,
in order to stimulate grid clients to submit packages. The here described
states will be represented to the clients in upcoming developments.

These states are specific for every compute element (CE) and communicated
between the Janitor and the Execution Service. Table~\ref{tab:states} shows all
possible states, while Figure~\ref{fig:RE_states}
displays the transitions between the states that a Runtime Environment may
be in during its life time at a particular CE. \task{?at a particular CE?}

@@

Upon presentation of a the package name to a Catalog, from which details about
the package are retrieved, a CE may
classify a package to be \texttt{INSTALLABLE} if all the dependencies
are installable or already \texttt{INSTALLED}. The installation can be
performed manually (\texttt{INSTALLING}/{\bf m}) or in an automated fashion
(\ldots/{\bf a}). Should the installation process return an error, then
the installation has \texttt{FAILED}. Once the installation succeeded,
the installed package is validated for its correctness. Should that
process fail, then the package's state it is said to be \texttt{BROKEN}.

@@

The Janitor manages the states that the runtime environments at a particular
compute element are in. However, it is also most important for the Janitor
to be aware of the jobs that depend on the installtion of a RE. REs still in use
should not be removed until the respective job has completed its computations.
The installation or removal of REs by the Janitor is perceived as a mere
consequene of jobs demanding a RE or not, thus, the communcation between
the job-manager AREX and the Janitor will be performed on that 'job level'.

The Janitor has two states for jobs: \texttt{PREPARED} and \texttt{INITIALIZED}. After a job has been succesfully registered 
in the Janitor, its state will be set to \texttt{PREPARED}. Invalid jobs are not cached. 
After the Janitor is requested to deploy the runtime environment, the state of the job will change to \texttt{INITIALIZED}. If 
an unforeseen exception occures during that process, the Janitor will drop the job from its database and set the affected runtime
environments to the state \texttt{FAILED}.

\section{Integration with AREX}

@@

In the current implementation the Janitor perl scripts must be executed as root. In order to execute Janitor with the required permissions,
a setuid wrapper has been prepared. During the installation \task{not done yet} the mode of rjanitor will be set to \texttt{u+s},
i.e. the sticky bit is set to make the binary suid root. \task{According to Daniel: ``A future version of the Janitor will get rid of the suid root helper.''}\\

Thus, Janitor is not executed directly. Instead, a dynamic link will be created in \task{libexec or sbin} which is pointing to
the wrapper.\\

@@

\subsection{What happens during installation}

Register
\begin{enumerate}
 \item Look up Catalog and find corresponding runtime environments.
 \item Check if dependcies are available (installed or installable)
 \item Store job in \textit{registrationdir}
\end{enumerate}

Deploy
\begin{enumerate}
 \item Load job out of \textit{registrationdir}
 \item Look up Catalog and find corresponding runtime environments.
 \item Check if dependcies are available (installed or installable)
 \item Download RTEs into the \textit{downloaddir}
 \item $\lbrack$ to be continued or skipped $\rbrack$
\end{enumerate}

@@

Security is a major concern for grid systems. Any additional feature
and especially an automatic software
installation inheritently introduces security threats. This section
addresses those and describes the available solutions to limit
security risks.

In the current installation, every user authorised to execute a job is
also authorised to install a REs. Restrictions are
only imposed on the set of dynamic REs that are available for installation.
Restrictions are imposed by the site admininistrators on the descriptions
that are given by the Catalog that is offering the package. These
descriptions may explicitly mention dynamic REs' names, e.g. a regular expression on
these, or refer to tags of packages that categorise these. However,
the core of these controls lies with the maintainers of the Catalog,
who needs to be trusted.

@@

\begin{itemize}
\item Dynamic REs are not distinguished between {\em installed} and
  {\em installable} in the information system.
\item No status information is given at the time a dynamic RE is installed.
\end{itemize}

Malevolent regular users with respective training in using system exploits
to gain root access are likely to find security holes by regularly
submitted scripts.  The authentication and authentification of users,
together with respective logging, is the major defense against such
attacks. What is consequently left to be protected against are unwanted
side-effects by the installation of software.

@@

The installation of packages from the Debian distribution (or other
packages of mainstream Linux distributions) is seeked to reduce the
complexity and burden in the maintenance for dynamic REs. In the
current implementation, Debian packages may be installed only by their
transformation into tar files. With the advent of the interface for
the virtualisation of the grid infrastructure, it is anticipated to
work with native packages of the Debian Linux distribution. The reuse
of packages that passed many eyeballs - as it is the case with packages
from major Linux distribution - security is further increased or becomes
as high as with the operating system underneath virtual clients.

>>>>

 The main language for the implementation of the functionality of the
 dynamic runtime environment functionality is Perl. And it is solely required (exceptions are the
 integration with the Grid Manager and the Web server) for the Janitor. In
 the pre-web-service implementation the Catalog remains a static web page.
 The Perl code is split into multiple modules as depicted in Figure~\ref{fig:janitorDependencies}. 
 The modules can be separated into two
 functional groups. One addresses the retrieval of information from
 the Catalog RDF file in the left major branch of the figure. The other
 addresses the process of fetching and installing the packages.

 \begin{landscape}
\begin{figure}[!h]
\vspace{4cm}
 \begin{center} 
    % \resizebox{24cm}{!}{\input{images/dependencies}}
    \mycaption{Modules of the Janitor and their dependencies}{}
 \end{center} 
\vfill
 \label{fig:janitorDependencies} 
\end{figure}
 \end{landscape}

@@

\section{Runtime environments states}
A main motivation for the managed, manual initiation of dynamic RE
installation is the subsequent manual verification of the installed
packages -- prior to their use in production.
With an automation of the installation, the verification of
that process shall be performed externally to that process.  At this
time, only the automation of the installation has been implemented.
To reflect the progress the external verification has made, REs
are said to be in states. The current implementation lists
installable REs aside the installed REs in the grid information system,
in order to stimulate grid clients to submit packages. The here described
states will be represented to the clients in upcoming developments.

These states are specific for every CE and communicated
between the Janitor and the Execution Service. Table~\ref{tab:states} shows all
possible states, while Figure~\ref{fig:RE_states}
displays the transitions between the states that a Runtime Environment may
be in during its life time at a particular CE. \task{?at a particular CE?}

@@

Upon presentation of a package to a Catalog, a CE may
classify a package to be \texttt{INSTALLABLE} if all the dependencies
are installable or already \texttt{INSTALLED}. The installation can be
performed manually (\texttt{INSTALLING}/{\bf m}) or in an automated fashion
(\ldots/{\bf a}). Should the installation process return an error, then
the installation has \texttt{FAILED}. Once the installation succeeded,
the installed package is validated for its correctness. Should that
process fail, then the package's state it is said to be \texttt{BROKEN}.

@@

Janitor has two states for jobs: \texttt{PREPARED} and \texttt{INITIALIZED}. After a job has been succesfully registered 
in Janitor, its state will be set to \texttt{PREPARED}. Invalid jobs are not cached in Janitor. 
After Janitor is requested to deploy the runtime environment, the state of the job will change to \texttt{INITIALIZED}. If 
an unforeseen exception occures during that process, Janitor will drop the job from its database and set the affected runtime
environments to the state \texttt{FAILED}.

\section{Integration into AREX}

@@

Currently the Janitor perl scripts must be executed as root. In order to execute Janitor with the required permissions,
a setuid wrapper has been written. During the installation \task{not done yet} the mode of rjanitor will be set to \texttt{u+s}
to make the binary suid root. \task{According to Daniel: ``A future version of the Janitor will get rid of the suid root helper.''}\\

Thus, Janitor is not executed directly. Instead a dynamic link will be created in \task{libexec or sbin} which is pointing to
the wrapper.\\

@@

@@

Security is a major concern for the grid systems. An automatic software
installation inheritently introduces security threats. This section
addresses those and describes the available solutions to limit
security risks.

In the current installation, every user authorised to execute a job is
also authorised to install a REs. Restrictions are
only imposed on the set of DRE that are available for installation.
Restrictions are imposed by the site admininistrators on the descriptions
that are given by the Catalog that is offering the package. These
descriptions may explicitly mention DRE names, a regular expression on
these, or refer to tags of packages that categorise these. However,
the core of these controls lie with the maintainers of the Catalog,
who needs to be trusted.

@@

\begin{itemize}
\item DREs are not distinguished between {\em installed} and
  {\em installable} in the information system.
\item No status information is given at the time an DRE is installed.
\end{itemize}

Malevolent regular users with respective training in using system exploits
to gain root access are likely to find security holes by regularly
submitted scripts.  The authentication and authentification of users,
together with respective logging, is the major defense against such
attacks. What is consequently left to be protected against are unwanted
sideeffects by the installation of software.

@@

The installation of packages from the Debian distribution (or other
packages of mainstream Linux distributions) is seeked to reduce the
complexity and burden in the maintenance for DRE. In the
current implementation Debian packages may be installed only by their
transformation into tar files. With the advent of the interface for
the virtualisation of the grid infrastructure, it is anticipated to
work with native packages of the Debian Linux distribution. The reuse
of packages that passed many eyeballs - as it is the case with packages
from major Linux distribution - security is further increased or becomes
as high as with the operating system underneath virtual clients.
